Reproducibility and plant pathology: where do we stand and a way forward

### 2021-06-16 analysis/paper/paper.Rmd
Grammar on line 101
Update cross-refs for figs and tables to work with word_document2 output

In biological sciences it is not always possible to use identical test material or perhaps the the time or resources are not available for full reproduction, _e.g._, field trials that span years and locations or complex glasshouse experiments.
Therefore we will follow Peng's [-@Peng2009] definition that provides clear guidelines a minimum standard of "reproducible research":

@@

Most articles did not make any computational methods available in any fashion with two, 0.007%, classing as "3", which was the highest score available (Figure \@ref(fig:article-scores)i).
Two hundred eighty-seven articles, 96%, were classed as "0" and nine articles, 0.03%, appeared to not use any computational methods.

More articles did a good job of using software that was reasonably available to anyone with 206 scoring "2" or "3" (Figure  \@ref(fig:article-scores)ii).

Software citations were also reasonably well reported in the literature with 120 scoring "2" where the main software was cited with version numbers and 57 scoring "3" where full citations including R packages or the SAS PROC used were cited where appropriate (Figure \@ref(fig:article-scores)iii).

However, data availability was mostly not available with 243 or 81.5% scoring "0" where the data was not available or mentioned in the article.
However, more articles, 29 or 9.7%, scored "3" than "1" or "2" combined, with two articles not producing shareable data (Figure  \@ref(fig:article-scores)iv).

The most frequently cited software was SAS.
The remainder were mostly specialised software except for Microsoft Excel, tied for 9th (Table \@ref(tab:su)).

@@

```{r journals, tab.id="journals", tab.cap="Journal titles selected for inclusion and the number (n) of articles from each journal that were evaluated.", echo=FALSE, message=FALSE, warning=FALSE}
library("dplyr")
library("janitor")
library("flextable")
library("officer")

import_notes() %>%
  tabyl(journal) %>%
  select(-percent) %>%
  rename(Journal = journal) %>% 
  flextable() %>%
  colformat_num(na_str = "NA") %>% 
  autofit()
```

@@

```{r software_used, tab.id="su", tab.cap="The top ten sofware programs that were found to be used in the papers that were surveyed.", echo=FALSE, message=FALSE, warning=FALSE}
library("tidyr")

rrpp_software <-
  import_notes() %>%
  transform(software_used = strsplit(software_used, ",")) %>%
  unnest(software_used) %>% 
  mutate(software_used = trimws(software_used)) %>% 
  mutate(software_used = toupper(software_used)) # convert all to uppercase to standardise

rrpp_software %>% 
  tabyl(software_used) %>%
  adorn_pct_formatting() %>% 
  select(c(-percent, -valid_percent)) %>% 
  rename(`Software Used` = software_used) %>% 
  arrange(-n) %>% 
  slice(seq_len(14)) %>% 
  drop_na() %>% 
  flextable() %>%
  autofit()
```

@@

```{r scoring, tab.id="scoring", tab.cap ="Articles were scored in four areas for reproducibility on a scale of '0' to '3' by four independent assessors to rank articles in each of four areas, 'Computational methods availability', 'Data availability', 'Software availability' and 'Software citation'.", echo=FALSE, message=FALSE, warning=FALSE}
scoring_criteria %>% 
  rename(`Index Value` = Index.Value) %>% 
  flextable() %>% 
  colformat_num(na_str = "NA") %>% 
  set_table_properties(layout = "autofit")
```

@@

```{r reproducibility_score, tab.id="reproducibility", tab.cap="Tally of reproducibility scores where '0' is not reproducible and '12' is fully reproducible, e.g., all software cited, software used are freely available to anyone, computational methods (statistical analysis or other computational code) are freely available for download and use, and raw data are available for download and use.", echo=FALSE, message=FALSE, warning=FALSE}
import_notes() %>%
  tabyl(reproducibility_score) %>%
  adorn_pct_formatting(digits = 1) %>%
  rename(Score = reproducibility_score, Percent = percent) %>%
  mutate(Score = round(Score, digits = 1)) %>% 
  flextable() %>%
  autofit()
```

@@

<!-- @Emerson, can you please have a look at this caption, if I'm not mistaken, this is your diagram design. Can you confirm that what I've written is accurate and describes the workflow? -->

```{r workflow-dia, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='An example open and reproducible research workflow. Starting with the question, determine the methodology, describe it, make it available and cite it. Data are used in analysis and any binary files or code are made available as supplements to the manuscript. Source code and raw data are made available in a public repository, preferably with version control for tracking changes through time and a DOI for final released products.', cache=FALSE, fig.env='figure'}
library("DiagrammeR")
library("DiagrammeRsvg")
library("rsvg")

@@

```{r article-scores, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Aggregated article scores for each of the four categories, (i), computational methods availability 0 - not available or not mentioned to 3 - freely available online to anonymous users. (ii) Shows 'Software availability', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Expensive proprietary software that only institutions would typically purchase'; '2' was 'Used proprietary software that most individuals can afford (e.g., Excel) or SAS that is now available free for use with a university login'; and '3' was 'Used entirely open source and free software (e.g., R, Julia, Python)'; 'NA' indicates that no specialised software was necessary or used. (iii) Displays 'Software citated', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Software mentioned by name only'; '2' was 'Software cited with version number'; and '3' was 'All software components (SAS PROCs, R, Julia or Python packages, etc.) cited'; 'NA' indicates that no specialised software was necessary or used. (iv) Shows 'Data availability', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Available upon request to the author; '2' was 'Online, but inconvenient or non-permanent (e.g., login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)'; and '3' was 'Freely available online to anonymous users for foreseeable future (e.g., archived using Zenodo, dataverse or university library or some other proper archiving system)'; 'NA' indicates that no data were generated, _e.g._ a methods paper.", cache=FALSE, fig.env='figure'}
library("ggplot2")
library("patchwork")
rrpp <- import_notes()
theme_set(theme_classic())
i <- ggplot(rrpp, aes(y = as.factor(comp_mthds_avail))) +
  geom_bar() +
  ggtitle("Computational Methods Availability") +
  ylab("Score") +
  xlab(NULL)

>>>>

In biological sciences it is not always possible to use identical test material or perhaps the the time or resources are not available for full reproduction, _e.g._, field trials that span years and locations or complex glasshouse experiments.
Therefore, we will follow Peng's [-@Peng2009] definition that provides clear guidelines a minimum standard of "reproducible research":

@@

Most articles did not make any computational methods available in any fashion with two, 0.007%, classing as "3", which was the highest score available (Figure \@ref(fig:articleScores)i).
Two hundred eighty-seven articles, 96%, were classed as "0" and nine articles, 0.03%, appeared to not use any computational methods.

More articles did a good job of using software that was reasonably available to anyone with 206 scoring "2" or "3" (Figure  \@ref(fig:articleScores)ii).

Software citations were also reasonably well reported in the literature with 120 scoring "2" where the main software was cited with version numbers and 57 scoring "3" where full citations including R packages or the SAS PROC used were cited where appropriate (Figure \@ref(fig:articleScores)iii).

However, data availability was mostly not available with 243 or 81.5% scoring "0" where the data was not available or mentioned in the article.
However, more articles, 29 or 9.7%, scored "3" than "1" or "2" combined, with two articles not producing shareable data (Figure \@ref(fig:articleScores)iv).

The most frequently cited software was SAS.
The remainder were mostly specialised software except for Microsoft Excel, tied for 9th (Table \@ref(tab:software-used)).

@@

```{r journals, echo=FALSE, message=FALSE, warning=FALSE}
import_notes() %>%
  tabyl(journal) %>%
  select(-percent) %>%
  rename(Journal = journal) %>%
  flextable() %>%
  colformat_num(na_str = "NA") %>%
  autofit() %>%
  set_caption(
    "Journal titles selected for inclusion and the number (n) of articles from each journal that were evaluated."
  )
```

@@

```{r software-used, echo=FALSE, message=FALSE, warning=FALSE}
rrpp_software <-
  import_notes() %>%
  transform(software_used = strsplit(software_used, ",")) %>%
  unnest(software_used) %>%
  mutate(software_used = trimws(software_used)) %>%
  mutate(software_used = toupper(software_used)) # convert all to uppercase to standardise

rrpp_software %>%
  tabyl(software_used) %>%
  adorn_pct_formatting() %>%
  select(c(-percent, -valid_percent)) %>%
  rename(`Software Used` = software_used) %>%
  arrange(-n) %>%
  slice(seq_len(14)) %>%
  drop_na() %>%
  flextable() %>%
  autofit() %>%
  set_caption(
    "The top ten sofware programs that were found to be used in the papers that were surveyed."
  )
```

@@

```{r scoring, echo=FALSE, message=FALSE, warning=FALSE}
scoring_criteria %>%
  rename(`Index Value` = Index.Value) %>%
  flextable() %>%
  colformat_num(na_str = "NA") %>%
  set_table_properties(layout = "autofit") %>%
  set_caption(
    "Articles were scored in four areas for reproducibility on a scale of '0' to '3' by four independent assessors to rank articles in each of four areas, 'Computational methods availability', 'Data availability', 'Software availability' and 'Software citation'.")
```

@@

```{r reproducibility_score, echo=FALSE, message=FALSE, warning=FALSE}
import_notes() %>%
  tabyl(reproducibility_score) %>%
  adorn_pct_formatting(digits = 1) %>%
  rename(Score = reproducibility_score, Percent = percent) %>%
  mutate(Score = round(Score, digits = 1)) %>%
  flextable() %>%
  autofit() %>%
  set_caption(
    "(\\#reproducibility) Tally of reproducibility scores where '0' is not reproducible and '12' is fully reproducible, e.g., all software cited, software used are freely available to anyone, computational methods (statistical analysis or other computational code) are freely available for download and use, and raw data are available for download and use."
  )
```

@@

```{r workflow-dia, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='An example open and reproducible research workflow. Starting with the question, determine the methodology, describe it, make it available and cite it. Data are used in analysis and any binary files or code are made available as supplements to the manuscript. Source code and raw data are made available in a public repository, preferably with version control for tracking changes through time and a DOI for final released products.', cache=FALSE, fig.env='figure'}

@@

```{r articleScores, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Aggregated article scores for each of the four categories, (i), computational methods availability 0 - not available or not mentioned to 3 - freely available online to anonymous users. (ii) Shows 'Software availability', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Expensive proprietary software that only institutions would typically purchase'; '2' was 'Used proprietary software that most individuals can afford (e.g., Excel) or SAS that is now available free for use with a university login'; and '3' was 'Used entirely open source and free software (e.g., R, Julia, Python)'; 'NA' indicates that no specialised software was necessary or used. (iii) Displays 'Software citated', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Software mentioned by name only'; '2' was 'Software cited with version number'; and '3' was 'All software components (SAS PROCs, R, Julia or Python packages, etc.) cited'; 'NA' indicates that no specialised software was necessary or used. (iv) Shows 'Data availability', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Available upon request to the author; '2' was 'Online, but inconvenient or non-permanent (e.g., login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)'; and '3' was 'Freely available online to anonymous users for foreseeable future (e.g., archived using Zenodo, dataverse or university library or some other proper archiving system)'; 'NA' indicates that no data were generated, e.g. a methods paper.", cache=FALSE, fig.env='figure'}

rrpp <- import_notes()
theme_set(theme_classic())
i <- ggplot(rrpp, aes(y = as.factor(comp_mthds_avail))) +
  geom_bar() +
  ggtitle("Computational Methods Availability") +
  ylab("Score") +
  xlab(NULL)
